\include{./../latex/notes_style.tex}
%--------------------------------------------------------------------------------------------------------------------------------

\begin{document}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  EXERCISES  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



 % ------------------------------------------------------------------------------------------------------------------------------------------
\begin{exercise}[]
Generate simulated that that is similar to the 
\end{exercise}


 % ------------------------------------------------------------------------------------------------------------------------------------------
\begin{exercise}[Divide by four rule]
Test the divide by for rule on the support for same-sex marriage data set by performing a linear regression on the binned data and comparing the coefficients to the logistic regression output. Clearly explain why this tests the divide by four rule.
\end{exercise}


 % ------------------------------------------------------------------------------------------------------------------------------------------
\begin{exercise}

\begin{Verbatim}
data = pd.read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv",encoding = "ISO-8859-1")
data["victim_age"] = pd.to_numeric(data["victim_age"],errors="coerce")

disp_new = []
disp = data["disposition"].values
for d in disp:
  if d == "Closed by arrest":
    disp_new.append(1)
  else:
    disp_new.append(0)
data["arrest"] = disp_new
data = data.dropna()
\end{Verbatim}
\end{exercise}


 % ------------------------------------------------------------------------------------------------------------------------------------------
\begin{exercise}[Water contamination model]

Suppose we have data consisting of the longitude ($X_1$) and latitude ($X_2$) of water samples along with whether or not the sample tested possitive for a toxin ($Y$). We believe that their has been a leak and want to identify the region within which there is more than a $1\%$ chance the water is contiminated.

How do we frame this problem as a logistic regression? We need to introduce features so that the contours of equal $q$ (the chance to be contaminated as a function of $X_1$ and $X_2$ are ellipsis. Recall the equation is
\begin{equation}
W = a_1X_1 + a_2X_2 + a_3X_1^2 + a_4X_2^2 + a_5X_1X_2
\end{equation}
The function $W(X_1,X_2)$ is constant along the curves.

This means that $h(W)$ is constant along these curves, where $h$ is the logistic function. It other words, $h$ will be largest at the center and decay.


\end{exercise}






\end{document}